---
title: "The Birds of Corhaven"
author: "Evelyn Yarzebinski & Kevin Soo"
output:
  pagedown::html_paged:
    toc: true
    toc_title: false
    number_sections: false
    chapter_name: false
---
```{r message=FALSE, echo=FALSE}
# Load libraries
library(tidyverse)
library(glue)
library(pagedown)

# Load data
df <- read_csv("tmp/corhaven_unique_birds_season_2022-02-08.csv",
               show_col_types = F) |> 
  select(-url, -photo, -description, -percent) |> 
  mutate(typicalSeason = as_factor(typicalSeason) |> 
           fct_relevel("Advent", "Christmas", "Epiphany", "Lent", "Easter", "Ordinary Time")) |> 
  group_by(birdFamily, comName, speciesCode) |> 
  summarize(typicalSeason = str_c(typicalSeason, collapse = ", "))

ebird <- read_csv("ebird_data.csv",
                  show_col_type = F) |> 
  mutate(photo_path = glue("photos/{speciesCode}.png"))

fam <- read_csv("bird-family-reflections.csv",
                show_col_types = F)

# Join dataframes
df <- df |> 
  left_join(ebird) 
```

##  {.page-break-before}

To Micah. May your early, wide-eyed wonder at the outdoors grow to include birds one day. -e

## Preface {.page-break-before}

We had the privilege of participating in the Coracle Fellowship in 2021. This field guide is our final project, borne out of the love, care, and reflection stimulated by that community.

Evelyn came up with the idea for this field guide after encountering some difficulties using a field guide as a novice birder. What if field guides could be more localized? Evelyn and Kevin, both data scientists, discussed how to bring this idea to light. We collaborated on writing code to query `eBird.org`, a citizen scientist website where birders can report their bird sightings and contribute to the overall understanding of birds and their behaviors. Our code downloaded all recorded observations of birds within a 5 kilometer radius of Corhaven in Quickburg, VA within the last 10 years. Evelyn wrote code to summarize and determine when each bird species tended to be observed near Corhaven in each season of the Christian liturgical calendar. Kevin wrote code to pull all of this together into this book.

Note that other than the code and reflections that underlie this guide, all other content - including bird descriptions and photos - comes from various sources, including Cornell University's eBird.org, National Geographic, and PBS. Birding, like Christianity, has a long history with many helpful contributions and perspectives from many individuals along the way. While innovation can be useful, it is often encouraged to discover and reference what others who have gone before have toiled to produce.

-e

# Part 1: On birding {.page-break-before}

## Introduction {.page-break-before}

It is no stretch to say that birding saved our lives. In early 2020, we were struggling through a season of personal darkness, which was compounded by being shut off from much of the outside world due to the Covid-19 pandemic. We spent many days watching the passage of time from the our dining room window, unable to see beyond our present reality in which we failed to flourish. 

It was around spring that we first started to notice the appearance (and then reappearance) of different kinds of birds through that same window. What were formerly "birds" to us became individual creatures. Different species became associated with distinct personalities, colors, and sounds. We bought a field guide to help us identify them, which transformed them in our eyes. They were *named*, they were *known*, and because they evoked in us a hint of the joy one feels when we realize someone has become familiar, we saw that they were *good.* 

The birds drew us outside of our home and ourselves. They called to us and gave us a reason to look and venture outwards. They gave structure to our days and weeks as we learned of their rhythms. Finally they became, to our surprise, reminders of God's kindness and presence with us through the seasons. Like God, birds come in and out of our view, but they are always around.

The significance of birds has a rich heritage within Christian tradition. Besides being mentioned and playing critical symbolic roles within the Bible, they have been used as symbols in Christian art for centuries. There were numerous stories in the early days of the church of its saints and mystics befriending birds. More recently, the late John Stott was a lover of birds, writing *The Birds Our Teachers: Biblical Lessons from a Lifelong Bird Watcher*, which can be found in the Corhaven library.

To anyone with eyes to see and ears to hear, we believe the practice of birding has much to offer us in our spiritual formation. It cultivates a sense of wonder and love of creation and forms good habits and virtues in those who give themselves to the practice. It has the potential to draw humans into deeper communion with the living God, who rules over much more than the birds, but certainly not less.

-k

## A note on Corhaven {.page-break-before}

Coracle is a living and breathing community of people knitted together by the Spirit and the humble work of its stewards. Its people are joined by a shared desire for deep spiritual communion, which transcends church traditions and localities. Despite its transcendent quality, Coracle is also tied intimately to a physical space. If you are reading this at Corhaven, you will be familiar with how the veil between God's space and ours seems to be just a little bit thinner there. I suspect that through common use, consecration, and cultivation of the grounds and buildings of Corhaven, it has been molded into a shape that is well-suited to revealing God's presence. 

Taking a walk around Corhaven, you will likely notice the chickens patrolling the grounds (they are not included in this field guide). Look above on a sunny day, and you may see Turkey Vultures circling overhead. Venture through Corhaven's paths into the woods and you will likely hear birds calling and singing (though they may be hard to spot). Like human communities, birds are tied to physical space and time. Specifically, most of them are tied to certain spaces at certain times -- i.e. they are seasonal. You can only really notice this if you try birding in the same place over time. 

Corhaven, in Quicksburg VA, is located close enough to migration pathways on the East coast to host a variety of birds. In addition, its setting among farmland, forests, and the Appalachian Mountains make it a welcoming habitat. Closer to Corhaven itself, the lush woods and Holman's Creek on the back of the grounds offer plenty of hiding and feeding spots for birds. All these factors make Corhaven a challenging but potentially rewarding place for birding.

-k

## Birding as spiritual practice {.page-break-before}

Easter Meadowlark is the only Advent bird - waiting season!
Helps us keep track of time, being in tune with God
Creation, not just matter

-k

## Birding as virtue formation {.page-break-before}

In which we acquire virtue.

## A note on the guide {.page-break-before}

Unlike other bird field guides, one innovation we included here is each entry includes the Christian liturgical season(s) in which they most commonly occur. We found this to be a surprisingly fitting, and we hope you agree. 

Creating an authoritative hyper-localized bird guide is a Sisyphean task: as soon as a new version is ready,  there will be weather systems, climate change, or other environmental factors that will influence the birds you may see in that area. You may not see the birds you expect to see, and you may even see birds you did not expect to see. Therefore, this guide was not designed to be authoritative. This guide is a combination of birds that either we have personally observed in the Corhaven area, or that other have reported observing on eBird.org. Your own mileage may vary.

... add note about seasons

-e 

# Part 2: Field guide {.page-break-before}

```{r echo=FALSE, results='asis'}
fam_template <- "\\newpage ## %s {data-short-title=''}

-----------------

%s

"

species_template <- "\\newpage ### %s {.unlisted}

![](%s){width=60%%}

%s

*Typical seasons: %s*

"

# for (i in seq(nrow(input))) {
#   current <- input[i, ]
#   cat(sprintf(template, current$name, current$data, current$text))
# }

for (i in fam$birdFamily) {
    
    # Get current bird family
    fam_current <- i
    
    # Get all species within bird family
    fam_species <- df |>
      filter(birdFamily == fam_current)
    
    # Get bird family reflection
    fam_reflection <- fam |>
      filter(birdFamily == i) |>
      pull(reflection)
    
    # Print bird family section header
    cat(sprintf(fam_template, fam_current, fam_reflection))
    
    # For all species within family, print an entry
    for (j in 1:nrow(fam_species)) {
      cat(sprintf(species_template, 
                  fam_species$comName[j], 
                  fam_species$photo_path[j],
                  fam_species$description[j],
                  fam_species$typicalSeason[j]
                  )
          )
      }
}
```